Forecasting the Changing Dynamics of US Energy Product Exports to OECD Markets: The Role of the Russian-Ukraine Conflict and Future Scenarios

### 2023-05-18 report.Rmd
updated e

@@

#checking accuracy
accuracy(fit)

fit %>% select(additive_ETS) %>% report()
additive_ets<-fit %>% select(additive_ETS)
addi_fc<- additive_ets %>% forecast(h=36)
accuracy(additi)
addi_fc %>% 
  autoplot(
    crude_oil_bx) +
  labs(
    y = "Thousand Barrels",
    title = "Forecasts for U.S. Crude oil exports to OECD Europe") +
  guides(colour = guide_legend(title = "Forecast"))
```

@@

For the training dataset, the additive model had the lowest ME (Mean Error) of -0.036 while RMSE (Root Mean Squared Error), MAE (Mean Absolute Error) and most of the other metrics are relatively similar when comparing Additive and Random models which are two best performing ETS models..

Additionally, the additive model had a lower ACF1 when comparing to other two models.A lower ACF1 (Autocorrelation Function at Lag 1) indicates a better model fit.

Considering these metrics, the additive model demonstrated superior performance in terms of forecasting accuracy for the crude oil time series. Hence, we selected the additive model (ETS=AAA) as the most appropriate ETS model for our analysis.

@@

```{r applying, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
#ETS model selection by R
crude_fit <- crude_oil_bx %>%
  model(auto_crude_ets = ETS(Total_Exports))
#checking the accuracy

accuracy(crude_fit)
report(crude_fit)
report(additive_ets)

auto_crude_fc<-crude_fit %>% select(auto_crude_ets) %>% forecast(h=36)

auto_crude_fc %>% 
  autoplot(crude_oil_bx) +
  labs(
    y = "Thousand Barrels",
    title = "Forecasts for U.S. Crude oil exports to OECD Europe") +
  guides(colour = guide_legend(title = "Forecast"))

@@

After applying the ETS() function to choose an appropriate model for the crude oil time series, we obtained the selected ETS model as <ETS(A,A,N)>. To assess the accuracy of the selected model, we calculated various metrics using the accuracy() function. The results indicate that the chosen model performs reasonably well in terms of forecasting accuracy. The ME (Mean Error) is close to zero, indicating that, on average, the model's forecasts are unbiased. The RMSE (Root Mean Squared Error) is 8.7, which represents the average magnitude of the forecast errors. The MAE (Mean Absolute Error) is 6.64, which measures the average absolute difference between the model's forecasts and the actual values. Finally, the ACF1 (Autocorrelation of Forecast Errors) is 0.129, representing the correlation between successive forecast errors. Overall, the selected ETS model demonstrates satisfactory accuracy, aligning with the main features observed in the crude oil time series.

@@

```{r bash, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
#Plotting Residual
gg_tsresiduals(additive_ets)
gg_tsresiduals(crude_fit)
```

In context of white noise, in the case of the additive model and crude_fit, the significant values at lag 12 and lag 15 respectively suggest that there is some correlation or pattern present in the residuals at those specific lags.

Therefore, based on the ACF plot, it can be concluded that the residuals in both models do not exhibit the characteristics of white noise. The presence of significant values in the ACF plot indicates some level of autocorrelation or dependency between the residuals at certain lags, suggesting that there are systematic patterns or information that the models have not fully captured.
Yet to be started

5.4 Comparison of ETS model to the simple models

```{r Comparion of Simple Models with ETS Model, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}

combined_accuracy <- accuracy(crude_oil_fit) %>%
  arrange(.model) %>%
  select(.model, .type, RMSE, MAE, MAPE, MASE, RMSSE) %>%
  bind_rows(
    accuracy(additive_ets),
    accuracy(crude_fit)
  )

combined_accuracy

@@

Among the forecast models evaluated, the additive_ETS and auto_crude_ets models demonstrate significantly better performance compared to the Drift, Mean, Naive, and Seasonal Naive models. The additive_ets model exhibits lower values for RMSE, MAE, MAPE, and MASE, indicating higher accuracy in predicting future values. Additionally, the model has a lower RMSSE value, suggesting a better fit to the data. While the ME is close to zero, indicating a minor bias in the forecasts, the MPE is negative, implying a slight underestimation of actual values on average. Also, low ACF1 suggests a weak positive correlation in the residuals at Lag 1. Similarly, the auto_crude_ets model shows favorable results with comparable accuracy metrics. Overall, the additive_ets and auto_crude_ets models outperform the other models, making them more suitable choices for forecasting.

>>>>

#Reliance on Russian gas

colnames(raw_rel_rusoil)

oecd_relrus <- ts_rel_rusoil %>%
  filter(Country == "OECD Europe")
oecd_relrus

@@

#checking accuracy
accuracy(fit) 
report(fit)
additive_ets<-fit %>% select(additive_ETS)

addi_fc<- fit %>% select(additive_ETS) %>% forecast(h=36)

addi_fc %>% 
  autoplot(
    crude_oil_bx) +
  labs(
    y = "Thousand Barrels",
    title = "Forecasts for U.S. Crude oil exports to OECD Europe") +
  guides(colour = guide_legend(title = "Forecast"))
```

@@

For the training dataset, the additive model had the lowest ME (Mean Error) of -0.036, suggesting that it has a relatively small average difference between its predictions and the actual values. Also ACF1 only of 0.11, which suggests that the residuals are reasonably independent and do not exhibit strong patterns or dependencies.

Additionally, the additive model had the lowest AICc, AIS and BIC compared to two other models, indicating that it provides a better balance between model complexity and goodness of fit. The lower values of AICc, AIC, and BIC suggest that the additive model outperforms the other models in terms of explaining the data while considering the number of parameters. This implies that the additive model is a more parsimonious choice and provides a better fit to the observed data, making it a preferred option for this analysis.

Considering these metrics, the additive model demonstrated superior performance in terms of forecasting accuracy for the natural gas time series. Hence, we selected the additive model (ETS=AAA) as the most appropriate ETS model for our analysis.

@@

```{r applying, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
#ETS model selection by R
crude_fit <- crude_oil_bx %>%
  model(auto_crude_ets = ETS(Total_Exports))
#checking the accuracy
report(crude_fit)
report(additive_ets)
accuracy(crude_fit)
accuracy(additive_ets)

auto_crude_ets<-crude_fit %>% select(auto_crude_ets)
auto_crud<-auto_crude_ets %>% forecast(h=36)

auto_crud%>% 
  autoplot(
    crude_oil_bx) +
  labs(
    y = "Thousand Barrels",
    title = "Forecasts for U.S. Crude oil exports to OECD Europe") +
  guides(colour = guide_legend(title = "Forecast"))

@@

After applying the ETS() function to choose an appropriate model for the crude oil time series, we obtained the selected ETS model as <ETS(A,A,N)>. To assess the accuracy of the selected model, we calculated various metrics using the accuracy() function. The results indicate that the chosen model performs reasonably well in terms of forecasting accuracy. The ME (Mean Error) is close to zero, indicating that, on average, the model's forecasts are unbiased. The ME of the model is -0.037 which is relatively small. Also, the ACF1 (Autocorrelation of Forecast Errors) is 0.129, representing the correlation between successive forecast errors. Finally, AIC, AICc and BIC criterias are also small comapred to previous models, which indicated that model is better fit in this case. Overall, the selected ETS model demonstrates satisfactory accuracy, aligning with the main features observed in the crude oil time series.

@@

```{r Checking residual of ETS models, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
#Plotting Residual
gg_tsresiduals(additive_ets)
gg_tsresiduals(auto_crude_ets)

@@

In context of white noise, in the case of the additive_ets model and auto_crude_ets, the significant values at lag 12 and lag 15 respectively suggest that there is some correlation or pattern present in the residuals at those specific lags.

Therefore, based on the ACF plot, it can be concluded that the residuals in both models do not exhibit the characteristics of white noise. The presence of significant values in the ACF plot indicates some level of autocorrelation or dependency between the residuals at certain lags, suggesting that there are systematic patterns or information that the models have not fully captured.

5.4 Comparison of ETS model against simple models
```{r Comparison of ETS model against simple models}
accuracy(crude_oil_fit) %>% 
  arrange(.model) %>% 
  select(.model, .type, RMSE, MAE, MAPE, MASE, RMSSE)
accuracy(additive_ets)
accuracy(auto_crude_ets)

combined_accuracy <- accuracy(crude_oil_fit) %>%
  arrange(.model) %>%
  select(.model, .type, RMSE, MAE, MAPE, MASE, RMSSE) %>%
  bind_rows(
    accuracy(additive_ets),
    accuracy(auto_crude_ets)
  )
combined_accuracy
```
In conclusion, when comparing our ETS models to the simple models from the Toolbox chapter, we consistently observe superior performance by the ETS models across all accuracy measures. This highlights the significant enhancement in accuracy achieved through the Exponential Smoothing approach. The ETS models consistently outperform the simple models, affirming the effectiveness and value of employing Exponential Smoothing techniques in forecasting.
